Internal models of type theory and internal higher categories for the Univalent Foundations

### 2014-02-25 article-models-tt.tex
started removing some of the redundant rules; added extension on terms

Note that by introducing extension on the level of types and families,
we introduce $\Sigma$-types at a very early stage.
However, we need substitution to make this precise.
\begin{align}
& \inference{\jfam{\Gamma}{A}}{\jctx{\ctxext{\Gamma}{A}}}
& & \inference{\jctxeq{\Gamma}{\Delta}\qquad\jfameq{\Gamma}{A}{B}}{\jctxeq{\ctxext{\Gamma}{A}}{\ctxext{\Delta}{B}}}\\
& \inference{\jfam{\ctxext{\Gamma}{A}}{P}}{\jfam{\Gamma}{\ctxext{A}{P}}}
& & \inference{\jfameq{\Gamma}{A}{B}\qquad\jfameq{\ctxext{\Gamma}{A}}{P}{Q}}{\jfameq{\Gamma}{\ctxext{A}{P}}{\ctxext{B}{Q}}}
\end{align}

@@

\subsection{Weakening}
We first define weakening by a context $\Gamma$. Since weakening by $\Gamma$
should in principle be a functor, it acts on contexts, types and terms alike.
Note that it is because of the weakening $\ctxwk{\Gamma}{\Delta}$ of a context
$\Delta$ by $\Gamma$ that we can speak of context morphisms from $\Gamma$ to $\Delta$: they are the terms
of $\ctxwk{\Gamma}{\Delta}$.

In the following inference rules we assume $\jctx{\Gamma}$ and in the rules
asserting a judgmental equality we furthermore assume $\jctxeq{\Gamma}{\Gamma'}$.
\begin{align}
& \inference{\jctx{\Delta}}{\jfam{\Gamma}{\ctxwk{\Gamma}{\Delta}}} 
& & \inference{\jctxeq{\Delta}{\Delta'}}{\jfameq{\Gamma}{\ctxwk{\Gamma}{\Delta}}{\ctxwk{\Gamma'}{\Delta'}}}\\
& \inference{\jfam{\Delta}{B}}{\jfam{\ctxext{\Gamma}{\ctxwk{\Gamma}{\Delta}}}{\ctxwk{\Gamma}{B}}}
& & \inference{\jfameq{\Delta}{B}{B'}}{\jfameq{\ctxext{\Gamma}{\ctxwk{\Gamma}{\Delta}}}{\ctxwk{\Gamma}{B}}{\ctxwk{\Gamma'}{B'}}}\\
& \inference{\jterm{\Delta}{B}{y}}{\jterm{\ctxext{\Gamma}{\ctxwk{\Gamma}{\Delta}}}{\ctxwk{\Gamma}{B}}{\ctxwk{\Gamma}{y}}} 
& & \inference{\jtermeq{\Delta}{B}{y}{y'}}{\jtermeq{\ctxext{\Gamma}{\ctxwk{\Gamma}{\Delta}}}{\ctxwk{\Gamma}{B}}{\ctxwk{\Gamma}{y}}{\ctxwk{\Gamma'}{y'}}} 
\end{align}

A weakening operation is also defined for types. When $A$ and $B$ are both types
 in context $\Gamma$, the weakened type $\ctxwk{A}{B}$ in context $\ctxext{\Gamma}{A}$
 is the family which `doesn't really depend on $A$'. The terms of $\ctxwk{A}{B}$
 are the functions from $A$ to $B$. Likewise, the terms $\ctxwk{A}{y}$ are the
 constant maps at $y$ from $A$ to $B$, for $y:B$.

@@

@@

Weakening by the empty context $\emptyc$ leaves contexts, context families and terms
unchanged:
\begin{align}
& \inference{\jctx{\Gamma}}{\jctxeq{\ctxwk{\emptyc}{\Gamma}}{\Gamma}}\\
& \inference{\jfam{\Gamma}{A}}{\jfameq{\Gamma}{\ctxwk{\emptyc}{A}}{A}}\\
& \inference{\jterm{\Gamma}{A}{x}}{\jtermeq{\Gamma}{A}{\ctxwk{\emptyc}{x}}{x}}
\end{align}

Likewise, weakening by the empty family $\emptyf$ over a context $\Gamma$ leaves families, families over that and
terms of those unchanged:
\begin{align}
& \inference{\jfam{\Gamma}{A}}{\jfameq{\Gamma}{\ctxwk{\emptyf}{A}}{A}}\\
& \inference{\jfam{\ctxext{\Gamma}{A}}{P}}{\jfameq{\ctxext{\Gamma}{A}}{\ctxwk{\emptyf}{P}}{P}}\\
& \inference{\jterm{\ctxext{\Gamma}{A}}{P}{f}}{\jtermeq{\ctxext{\Gamma}{A}}{P}{\ctxwk{\emptyf}{f}}{f}}
\end{align}

The following rules assert the compatibility of extension with weakening: for
every context $\Gamma$ and every type $B$ in context $\Delta$ there is a
judgmental equality $\ctxwk{\Gamma}{\ctxext{\Delta}{B}}\jdeq\ctxext{\ctxwk{\Gamma}{\Delta}}
{\ctxwk{\Gamma}{B}}$. When thinking of terms of $\ctxwk{\Gamma}{\Delta}$ as
functions from $\Gamma$ to $\Delta$, this looks awfully lot like the type
theoretical axiom of choice, of which it seems absurd to ask to hold judgmentally.
But the rule should be there because in ordinary Martin-L\"of type theory
there is no syntax for weakening, i.e.~weakening is invisible. 
And thus both $\ctxwk{\Gamma}{\ctxext{\Delta}{B}}$ 
and $\ctxext{\ctxwk{\Gamma}{\Delta}}{\ctxwk{\Gamma}{B}}$ is the type
$\ctxext{\Delta}{B}$ in context $\Gamma$.

In the following inference rules we assume that $\jctx{\Gamma}$.

\begin{align}
& \inference
  {\jfam{\Delta}{B}}
  {\jfameq{\Gamma}{\ctxwk{\Gamma}{\ctxext{\Delta}{B}}}{\ctxext{\ctxwk{\Gamma}{\Delta}}{\ctxwk{\Gamma}{B}}}}\\
& \inference
  {\jfam{\Delta}{B}\qquad\jfam{\ctxext{\Delta}{B}}{Q}}
  {\jfameq
    {\ctxext{\Gamma}{\ctxwk{\Gamma}{\Delta}}}
    {\ctxwk{\Gamma}{\ctxext{B}{Q}}}
    {\ctxext{\ctxwk{\Gamma}{B}}{\ctxwk{\Gamma}{Q}}}}
\end{align}

In the following inference rules we assume that $\jfam{\Gamma}{A}$.

\begin{align}
& \inference
  {\jfam{\ctxext{\Gamma}{B}}{Q}}
  {\jfameq{\ctxext{\Gamma}{A}}{\ctxwk{A}{\ctxext{B}{Q}}}{\ctxext{\ctxwk{A}{B}}{\ctxwk{A}{Q}}}}\\
& \inference
  {\jfam{\ctxext{{\Gamma}{B}}{Q}}{R}}
  {\jfameq
    {\ctxext{{\Gamma}{A}}{\ctxwk{A}{B}}}
    {\ctxwk{A}{\ctxext{Q}{R}}}
    {\ctxext{\ctxwk{A}{Q}}{\ctxwk{A}{R}}}}
\end{align}

\subsubsection{Weakening is compatible with itself}\label{comp-ww}
We state judgmental equality rules expressing
that weakening is compatible with itself. These rules state that the following
diagram commutes given any two contexts $\Gamma$ and $\Delta$:
\begin{equation*}
\begin{tikzcd}[column sep=huge]
\jctx{\blank} \ar{r}{\greek{E}\mapsto\ctxwk{\Delta}{E}} \ar{d}[swap]{\greek{E}\mapsto\ctxwk{\Gamma}{E}} & \jfam{\Delta}{\blank} \ar{d}{B\mapsto\ctxwk{\Gamma}{B}}\\
\jfam{\Gamma}{\blank} \ar{r}[swap]{A\mapsto\ctxwk{{\Gamma}{\Delta}}{A}} & \jfam{\ctxext{\Gamma}{\ctxwk{\Gamma}{\Delta}}}{\blank}
\end{tikzcd}
\end{equation*}
There is also a version of this diagram in which all happens in a context. Thus,
we get two sets of inference rules. For weakening by contexts we get:
\begin{align}
& \inference{\jctx{\Gamma}\qquad\jctx{\Delta}\qquad\jctx{\greek{E}}}
          {\jfameq{\ctxext{\Gamma}{\ctxwk{\Gamma}{\Delta}}}{\ctxwk{\Gamma}{{\Delta}{\greek{E}}}}
            {\ctxwk{{\Gamma}{\Delta}}{{\Gamma}{\greek{E}}}}}\\
& \inference{\jctx{\Gamma}\qquad\jctx{\Delta}\qquad\jfam{\greek{E}}{C}}
          {\jfameq{\ctxext{{\Gamma}{\ctxwk{\Gamma}{\Delta}}}{\ctxwk{{\Gamma}{\Delta}}{{\Gamma}{\greek{E}}}}}
            {\ctxwk{\Gamma}{{\Delta}{C}}}
            {\ctxwk{{\Gamma}{\Delta}}{{\Gamma}{C}}}}\\
& \inference{\jctx{\Gamma}\qquad\jctx{\Delta}\qquad\jterm{\greek{E}}{C}{t}}
          {\jtermeq{\ctxext{{\Gamma}{\ctxwk{\Gamma}{\Delta}}}{\ctxwk{{\Gamma}{\Delta}}{{\Gamma}{\greek{E}}}}}
            {\ctxwk{{\Gamma}{\Delta}}{{\Gamma}{C}}}{\ctxwk{\Gamma}{{\Delta}{t}}}{\ctxwk{{\Gamma}{\Delta}}{{\Gamma}{t}}}}
\end{align}
For weakening by types we get:
\begin{align}
& \inference{\jfam{\Gamma}{A}\qquad\jfam{\Gamma}{B}\qquad\jfam{\Gamma}{C}}
          {\jfameq{\ctxext{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{{B}{C}}}
            {\ctxwk{{A}{B}}{{A}{C}}}}\\
& \inference{\jfam{\Gamma}{A}\qquad\jfam{\Gamma}{B}\qquad\jfam{\ctxext{\Gamma}{C}}{R}}
          {\jfameq{\ctxext{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{{A}{B}}{{A}{C}}}}
            {\ctxwk{A}{{B}{R}}}
            {\ctxwk{{A}{B}}{{A}{R}}}}\\
& \inference{\jfam{\Gamma}{A}\qquad\jfam{\Gamma}{B}\qquad\jterm{\Gamma}{R}{t}}
          {\jtermeq{\ctxext{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{{A}{B}}{{A}{C}}}}
            {\ctxwk{{A}{B}}{{A}{R}}}{\ctxwk{A}{{B}{t}}}{\ctxwk{{A}{B}}{{A}{t}}}}
\end{align}

@@

@@

In the following, we assume that $\jctx{\Gamma}$ and $\jfam{\Gamma}{A}$.

\begin{align}
& \inference{\jctx{\Delta}}
    {\jfameq{\ctxext{\Gamma}{A}}{\ctxwk{\ctxext{\Gamma}{A}}{\Delta}}{\ctxwk{A}{{\Gamma}{\Delta}}}}\\
& \inference{\jfam{\Delta}{B}}
    {\jfameq{\ctxext{{\Gamma}{A}}{\ctxwk{A}{{\Gamma}{\Delta}}}}{\ctxwk{\ctxext{\Gamma}{A}}{B}}{\ctxwk{A}{{\Gamma}{B}}}}\\
& \inference{\jterm{\Delta}{B}{g}}
    {\jtermeq{\ctxext{{\Gamma}{A}}{\ctxwk{A}{{\Gamma}{\Delta}}}}{\ctxwk{A}{{\Gamma}{B}}}{\ctxwk{\ctxext{\Gamma}{A}}{g}}{\ctxwk{A}{{\Gamma}{g}}}}
\end{align}

The rules for weakening by a type and a family over it are similar. We assume that
$\jfam{\Gamma}{A}$ and $\jfam{\ctxext{\Gamma}{A}}{P}$.

\begin{align}
& \inference{\jfam{\Gamma}{B}}
    {\jfameq{\ctxext{{\Gamma}{A}}{P}}{\ctxwk{\ctxext{A}{P}}{B}}{\ctxwk{P}{{A}{B}}}}\\
& \inference{\jfam{\ctxext{\Gamma}{B}}{Q}}
    {\jfameq{\ctxext{{{\Gamma}{A}}{P}}{\ctxwk{P}{{A}{B}}}}{\ctxwk{\ctxext{A}{P}}{Q}}{\ctxwk{P}{{A}{Q}}}}\\
& \inference{\jterm{\Gamma}{B}{g}}
    {\jtermeq{\ctxext{{\Gamma}{A}}{P}}{\ctxwk{P}{{A}{B}}}{\ctxwk{\ctxext{A}{P}}{g}}{\ctxwk{P}{{A}{g}}}}\\
& \inference{\jterm{\ctxext{\Gamma}{B}}{Q}{s}}
    {\jtermeq{\ctxext{{{\Gamma}{A}}{P}}{\ctxwk{P}{{A}{B}}}}{\ctxwk{P}{{A}{Q}}}{\ctxwk{\ctxext{A}{P}}{s}}{\ctxwk{P}{{A}{s}}}}
\end{align}

\subsection{Substitution}
Given a family $P$ over $A$ and a term $x$ of $A$, substitution gives a way to
consider the fiber $\subst{x}{P}$ of $P$ at $x$. Also, we get a way to evaluate
terms $f$ of $P$ at $x$. This will give us ways to compose functions too. In
this section, we shall first introduce the operations `substitution of a term $x$'
for types, terms and families. Then we shall explain how substitution interacts
with itself, extension and weakening.

@@

>>>>

\begin{align}
& \inference{\jfam{\Gamma}{A}}{\jctx{\ctxext{\Gamma}{A}}}
& & \inference{\jctxeq{\Gamma}{\Delta}\qquad\jfameq{\Gamma}{A}{B}}{\jctxeq{\ctxext{\Gamma}{A}}{\ctxext{\Delta}{B}}}\\
& \inference{\jfam{\ctxext{\Gamma}{A}}{P}}{\jfam{\Gamma}{\ctxext{A}{P}}}
& & \inference{\jfameq{\Gamma}{A}{B}\qquad\jfameq{\ctxext{\Gamma}{A}}{P}{Q}}{\jfameq{\Gamma}{\ctxext{A}{P}}{\ctxext{B}{Q}}}
\end{align}

@@

\subsection{The type theoretic operation of weakening}
When $A$ is a context family over a context $\Gamma$, we wish to define a weakening
operation $\ctxwk{A}{}$. The weakening operation acts on context families $B$ 
over $\Gamma$, terms thereof, context families over $B$ and terms thereof.
It weakens those, which means that it ``adds $A$ to the context''. The context
family $\ctxwk{A}{B}$ can be seen as the constant family $B$ over $\ctxext{\Gamma}{A}$.
Likewise, when $y$ is a term of $B$, the term $\ctxwk{A}{y}$ of $\ctxwk{A}{B}$
can be seen as the constant term with value $y$.

@@

We add rules asserting that a weakened type is again a type:

\begin{align}
& \inference{\jtype{\Gamma}{B}}{\jtype{\ctxext{\Gamma}{A}}{\ctxwk{A}{B}}}\\
& \inference{\jfam{\Gamma}{B}\quad\jtype{\ctxext{\Gamma}{B}}{Q}}{\jtype{\ctxext{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{Q}}}
\end{align}

@@

Weakening by the empty family $\emptyf$ over a context $\Gamma$ leaves families, 
their terms, families over those families and
terms of those unchanged:
\begin{align}
& \inference{\jfam{\Gamma}{B}}{\jfameq{\Gamma}{\ctxwk{\emptyf}{B}}{B}}\\
& \inference{\jterm{\Gamma}{B}{y}}{\jtermeq{\Gamma}{B}{\ctxwk{\emptyf}{y}}{y}}\\
& \inference{\jfam{\ctxext{\Gamma}{B}}{Q}}{\jfameq{\ctxext{\Gamma}{B}}{\ctxwk{\emptyf}{Q}}{Q}}\\
& \inference{\jterm{\ctxext{\Gamma}{B}}{Q}{g}}{\jtermeq{\ctxext{\Gamma}{B}}{Q}{\ctxwk{\emptyf}{g}}{g}}
\end{align}

@@

The following rules assert the compatibility of extension with weakening: for
every family $A$ over $\Gamma$ and every family $Q$ over $\ctxext{\Gamma}{B}$
there is a
judgmental equality $\ctxwk{A}{\ctxext{B}{Q}}\jdeq\ctxext{\ctxwk{A}{B}}
{\ctxwk{A}{Q}}$.

When thinking of terms of $\ctxwk{A}{B}$ as morphisms of families from $A$ to
$B$, this looks already like form of type theoretic choice. It is weaker in that
it is not stated with function types, yet it is stronger in that it states a
judgmental equality between two families. When one makes the weakening operation
notationally invisible -- as is in fact the usual practice in type theory -- the
following compatibility rules become completely obvious.

In the following inference rules we assume that $\jfam{\Gamma}{A}$.
\begin{align}
& \inference
  {\jfam{\ctxext{\Gamma}{B}}{Q}}
  {\jfameq{\ctxext{\Gamma}{A}}{\ctxwk{A}{\ctxext{B}{Q}}}{\ctxext{\ctxwk{A}{B}}{\ctxwk{A}{Q}}}}\\
& \inference
  {\jfam{\ctxext{{\Gamma}{B}}{Q}}{R}}
  {\jfameq
    {\ctxext{{\Gamma}{A}}{\ctxwk{A}{B}}}
    {\ctxwk{A}{\ctxext{Q}{R}}}
    {\ctxext{\ctxwk{A}{Q}}{\ctxwk{A}{R}}}}
\end{align}

\subsubsection{Weakening is compatible with itself}\label{comp-ww}
We state judgmental equality rules expressing
that weakening is compatible with itself. These rules state that the following
diagram commutes given any two families $A$ and $B$ in context $\Gamma$:
\begin{equation*}
\begin{tikzcd}[column sep=huge]
\jfam{\Gamma}{\blank} \ar{r}{C\mapsto\ctxwk{B}{C}} \ar{d}[swap]{C\mapsto\ctxwk{A}{C}} & \jfam{\ctxext{\Gamma}{B}}{\blank} \ar{d}{Q\mapsto\ctxwk{A}{Q}}\\
\jfam{\ctxext{\Gamma}{A}}{\blank} \ar{r}[swap]{P\mapsto\ctxwk{{A}{B}}{P}} & \jfam{\ctxext{{\Gamma}{A}}{\ctxwk{A}{B}}}{\blank}
\end{tikzcd}
\end{equation*}
Thus, we get the following set of inference rules:
\begin{align}
& \inference{\jfam{\Gamma}{A}\qquad\jfam{\Gamma}{B}\qquad\jfam{\Gamma}{C}}
          {\jfameq{\ctxext{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{{B}{C}}}
            {\ctxwk{{A}{B}}{{A}{C}}}}\\
& \inference{\jfam{\Gamma}{A}\qquad\jfam{\Gamma}{B}\qquad\jterm{\Gamma}{C}{z}}
          {\jtermeq{\ctxext{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{{A}{B}}{{A}{C}}}
            {\ctxwk{A}{{B}{z}}}
            {\ctxwk{{A}{B}}{{A}{z}}}}\label{comp-ww-inf2}\\
& \inference{\jfam{\Gamma}{A}\qquad\jfam{\Gamma}{B}\qquad\jfam{\ctxext{\Gamma}{C}}{R}}
          {\jfameq{\ctxext{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{{A}{B}}{{A}{C}}}}
            {\ctxwk{A}{{B}{R}}}
            {\ctxwk{{A}{B}}{{A}{R}}}}\\
& \inference{\jfam{\Gamma}{A}\qquad\jfam{\Gamma}{B}\qquad\jterm{\Gamma}{R}{t}}
          {\jtermeq{\ctxext{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{{A}{B}}{{A}{C}}}}
            {\ctxwk{{A}{B}}{{A}{R}}}{\ctxwk{A}{{B}{t}}}{\ctxwk{{A}{B}}{{A}{t}}}}
            \label{comp-ww-inf4}
\end{align}
Note that strictly speaking, we did not need to require the inference in
\autoref{comp-ww-inf2} since this follows from the inference in \autoref{comp-ww-inf4}
and the fact that the weakening of the empty family is again the empty family.

@@

In the following inference rules we assume that
$\jfam{\Gamma}{A}$ and $\jfam{\ctxext{\Gamma}{A}}{P}$. 
\begin{align}
& \inference{\jfam{\Gamma}{B}}
    {\jfameq{\ctxext{{\Gamma}{A}}{P}}{\ctxwk{\ctxext{A}{P}}{B}}{\ctxwk{P}{{A}{B}}}}\\
& \inference{\jterm{\Gamma}{B}{y}}
    {\jtermeq{\ctxext{{\Gamma}{A}}{P}}{\ctxwk{P}{{A}{B}}}{\ctxwk{\ctxext{A}{P}}{y}}{\ctxwk{P}{{A}{y}}}}\\
& \inference{\jfam{\ctxext{\Gamma}{B}}{Q}}
    {\jfameq{\ctxext{{{\Gamma}{A}}{P}}{\ctxwk{P}{{A}{B}}}}{\ctxwk{\ctxext{A}{P}}{Q}}{\ctxwk{P}{{A}{Q}}}}\\
& \inference{\jterm{\ctxext{\Gamma}{B}}{Q}{g}}
    {\jtermeq{\ctxext{{{\Gamma}{A}}{P}}{\ctxwk{P}{{A}{B}}}}{\ctxwk{P}{{A}{Q}}}{\ctxwk{\ctxext{A}{P}}{g}}{\ctxwk{P}{{A}{g}}}}
\end{align}

\subsection{The type theoretic operation of substitution}
Given a family $P$ over $A$ and a term $x$ of $A$, substitution gives a way to
consider the fiber $\subst{x}{P}$ of $P$ at $x$. Also, we get a way to evaluate
terms $f$ of $P$ at $x$. This will give us ways to compose functions too. In
this section, we shall first introduce the operations `substitution of a term $x$'
for types, terms and families. Then we shall explain how substitution interacts
with itself, extension and weakening.

@@

\subsubsection{Extension on terms}
When $\jterm{\Gamma}{A}{x}$ and $\jterm{\Gamma}{\subst{x}{P}}{u}$ are terms,
we obtain a term $\jterm{\Gamma}{\ctxext{A}{P}}{\ctxext{x}{u}}$. This operation
of weakening on terms is like the pairing operation. However, it is handled
much stricter for it has a judgmental inverse and it is subject to various compatibility
rules.
\begin{align}
& \inference{\jterm{\Gamma}{A}{x}\quad\jterm{\Gamma}{\subst{x}{P}}{u}}{\jterm{\Gamma}{\ctxext{A}{P}}{\ctxext{x}{u}}}\\
& \inference{\jterm{\Gamma}{\ctxext{A}{P}}{w}}{\jterm{\Gamma}{A}{\pts{w}}}\\
& \inference{\jterm{\Gamma}{\ctxext{A}{P}}{w}}{\jterm{\Gamma}{\subst{\pts{w}}{P}}{\edg{w}}}\\
& \inference{\jterm{\Gamma}{\ctxext{A}{P}}{w}}{\jtermeq{\Gamma}{\ctxext{A}{P}}{\ctxext{\pts{w}}{\edg{w}}}{w}}\\
& \inference{\jterm{\Gamma}{A}{x}\quad\jterm{\Gamma}{\subst{x}{P}}{u}}{\jtermeq{\Gamma}{A}{\pts{(\ctxext{x}{u})}}{x}}\\
& \inference{\jterm{\Gamma}{A}{x}\quad\jterm{\Gamma}{\subst{x}{P}}{u}}{\jtermeq{\Gamma}{\subst{x}{P}}{\edg{(\ctxext{x}{u})}}{u}}
\end{align}

We add the following compatibility rule for two consecutive term extensions:
\begin{equation}
\inference{\jterm{\Gamma}{A}{x}\quad\jterm{\Gamma}{\subst{x}{P}}{u}\quad\jterm{\Gamma}{\subst{\ctxext{x}{u}}{Q}}{v}}
{\jtermeq{\Gamma}{\ctxext{{A}{P}}{Q}}{\ctxext{x}{{u}{v}}}{\ctxext{{x}{u}}{v}}}
\end{equation}

When we weaken a term $\ctxext{y}{v}$ of $\ctxext{B}{Q}$ in context $\Gamma$ by
a type $A$, the term that we get is $\ctxext{\ctxwk{A}{y}}{\ctxwk{A}{v}}$.
\begin{align}
& \inference{\jterm{\Gamma}{B}{y}\quad\jterm{\Gamma}{\subst{y}{Q}}{v}}{\jtermeq{\ctxext{\Gamma}{A}}{\ctxwk{A}{\ctxext{B}{Q}}}{\ctxwk{A}{\ctxext{y}{v}}}{\ctxext{\ctxwk{A}{y}}{\ctxwk{A}{v}}}}\\
& \inference{\jterm{\ctxext{\Gamma}{B}}{Q}{g}\quad\jterm{\ctxext{\Gamma}{B}}{\subst{g}{R}}{t}}{\jtermeq{\ctxext{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{\ctxext{Q}{R}}}{\ctxwk{A}{\ctxext{g}{t}}}{\ctxext{\ctxwk{A}{g}}{\ctxwk{A}{t}}}}
\end{align}

When we substitute an extended term $\ctxext{f}{g}$ of $\ctxext{P}{Q}$ by a term
$x$ of $A$, the term that we get is $\ctxext{\subst{x}{f}}{\subst{x}{g}}$.
\begin{align}
& \inference{\jterm{\ctxext{\Gamma}{A}}{P}{f}\quad\jterm{\ctxext{\Gamma}{A}}{\subst{f}{Q}}{g}}{\jtermeq{\Gamma}{\ctxext{\subst{x}{P}}{\subst{x}{Q}}}{\subst{x}{\ctxext{f}{g}}}{\ctxext{\subst{x}{f}}{\subst{x}{g}}}}\\
& \inference{\jterm{\ctxext{{\Gamma}{A}}{P}}{Q}{g}\quad\jterm{\ctxext{{\Gamma}{A}}{P}}{\subst{g}{R}}{t}}{\jtermeq{\ctxext{\Gamma}{\subst{x}{P}}}{\ctxext{\subst{x}{Q}}{\subst{x}{R}}}{\subst{x}{\ctxext{g}{t}}}{\ctxext{\subst{x}{g}}{\subst{x}{t}}}}
\end{align}

Finally, when we substitute by an extended term we get an equal result as when we
substitute two consecutive times. This looks like the Curry-Howard isomorphism.
\begin{align}
& \inference{\jterm{\Gamma}{A}{x}\quad\jterm{\Gamma}{\subst{x}{P}}{u}\quad\jfam{\ctxext{{\Gamma}{A}}{P}}{Q}}
  {\jfameq{\Gamma}{\subst{\ctxext{x}{u}}{Q}}{\subst{u}{{x}{Q}}}}\\
& \inference{\jterm{\Gamma}{A}{x}\quad\jterm{\Gamma}{\subst{x}{P}}{u}\quad\jterm{\ctxext{{\Gamma}{A}}{P}}{Q}{g}}
  {\jtermeq{\Gamma}{\subst{u}{{x}{Q}}}{\subst{\ctxext{x}{u}}{g}}{\subst{u}{{x}{g}}}}\\
& \inference{\jterm{\Gamma}{A}{x}\quad\jterm{\Gamma}{\subst{x}{P}}{u}\quad\jfam{\ctxext{{{\Gamma}{A}}{P}}{Q}}{R}}
  {\jfameq{\ctxext{\Gamma}{\subst{u}{{x}{Q}}}}{\subst{\ctxext{x}{u}}{R}}{\subst{u}{{x}{R}}}}\\
& \inference{\jterm{\Gamma}{A}{x}\quad\jterm{\Gamma}{\subst{x}{P}}{u}\quad\jterm{\ctxext{{{\Gamma}{A}}{P}}{Q}}{R}{t}}
  {\jtermeq{\ctxext{\Gamma}{\subst{u}{{x}{Q}}}}{\subst{u}{{x}{R}}}{\subst{\ctxext{x}{u}}{t}}{\subst{u}{{x}{t}}}}
\end{align}
